
\documentclass[hidelinks,12pt]{article}
\usepackage{lastpage}
\usepackage{fancyhdr}
\fancyfoot{}
\cfoot{\thepage{}~of~\pageref{LastPage}}
\usepackage{caption}
\usepackage[pdftex]{graphicx}
\usepackage{epstopdf}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amstext}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{breqn}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage{bigints}
\usepackage{color}
%\usepackage{parskip}
\usepackage[letterpaper]{geometry}
\usepackage{geometry}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{keyval}
\usepackage{ifthen}
%\usepackage[american]{babel}
\usepackage{etoolbox}
\usepackage[all]{nowidow}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{comment}
\usepackage{placeins}
\usepackage{upgreek}
%\usepackage{natbib}
\usepackage{array}
\usepackage{adjustbox}
\usepackage[demo,abs]{overpic}
\usepackage{booktabs}

\definecolor{cadmiumgreen}{rgb}{0.0, 0.42, 0.24}
%% TO CONVERT TO TIMES NEW ROMAN
%\usepackage{mathptmx}
% convert table numering to roman
%\renewcommand*\thetable{\Roman{table}}

\usepackage[backend = biber, authordate, uniquename=false, uniquelist=false, url = false, doi=false, isbn=false]{biblatex-chicago}
\externaldocument{bpw_water_paper_appendix}

\addbibresource{bpw_water.bib}


\geometry{
top = 1in,            % <-- you want to adjust this
inner = 1in,
outer = 1in,
bottom = 1in,
%headheight = 3ex,       % <-- and this
%headsep = 2ex,          % <-- and this
}

\title{\vspace{-10mm} {Groundwater, energy, and crop choice}
}
\date{\today\\ \vspace{0.3cm} %\textsc{Preliminary and Incomplete \\ Not for citation or circulation}
}
\author{
Fiona Burlig, Louis Preonas, and Matt Woerman\thanks{\scriptsize Burlig: University of Chicago Harris School of Public Policy and NBER. Keller Center, 1307 E. 60th Street, Chicago, IL 60637. Email: \url{burlig@uchicago.edu}. Preonas: Department of Agricultural and Resource Economics, University of Maryland. 2200 Symons Hall, 7998 Regents Drive, College Park, MD 20742. Email: \url{lpreonas@umd.edu}. Woerman: Department of Resource Economics, University of Massachusetts Amherst. Stockbridge Hall, 80 Campus Center Way, Amherst, MA 01003. Email: \url{mwoerman@umass.edu}. 
%
We thank Nick Hagerty for graciously sharing data with us, and thank Amy Ando, Andrew Ayers, Maximilian Auffhammer, Kathy Baylis, Ellen Bruno, Karen Clay, Michael Greenstone, Koichiro Ito, Katrina Jessoe, Erik Lichtenberg, Dave McLaughlin, Kyle Meng, Mar Reguant, Paul Scott, Joe Shapiro, Andrew Stevens, Catherine Wolfram, and seminar participants at UC Energy Camp, Mississippi State, UC Davis, University of Illinois, ITAM, HKUST, University of Chicago, USC, the ASSA Annual Meetings, the AERE Summer Conference, the WEAI annual meetings, and UC Berkeley's POWER Conference for helpful comments and suggestions.
%
Chinmay Lohani, Anna Schmidt, Chen Sui, Yixin Sun, and Xinyi Wang provided excellent research assistance.
%
Karen Notsund provided invaluable support in obtaining data. 
%
We received generous funding in support of this project from the Giannini Foundation of Agricultural Economics, and the Sloan Foundation (via the E2e Project), and the Becker Friedman Institute for Economics.
%
All remaining errors are our own.}}
\begin{document}

\setlength{\parindent}{0.9cm}
\maketitle
\thispagestyle{empty}
\normalsize
%\vspace{-2mm}
%\thispagestyle{empty}
\begin{abstract}
%\vspace{-5mm} 
\input{bpw_water_paper_abstract}

\noindent \textbf{Keywords:} groundwater, agriculture, electricity \\
\noindent \textbf{JEL Codes:} Q15, Q25, Q41 \\
 

\end{abstract}
\vspace{5mm}

\newpage
\clearpage
\setcounter{page}{1}
\begin{refsection}
\begin{spacing}{1.5} {

\section{Introduction}
\label{sec:intro}
\input{bpw_water_paper_sec1}

\section{Background}
\label{sec:background}
\input{bpw_water_paper_sec2}


\section{Data}
\label{sec:data}
\input{bpw_water_paper_sec3}


\section{Empirical strategy}
\label{sec:empirics}
\input{bpw_water_paper_sec4}


\section{Results}
\label{sec:results}
\input{bpw_water_paper_sec5}


\section{Economics of groundwater demand response}
\label{sec:mechanisms}
\input{bpw_water_paper_sec6}

\section{Conclusion}
\label{sec:conclusion}
\input{bpw_water_paper_sec7}
}
\end{spacing}


\section*{References}
\printbibliography[heading=none]
\newpage
\end{refsection}
\FloatBarrier
\section*{Figures and Tables}
$\phantom{SPACER}$

$\phantom{SPACER}$

$\phantom{SPACER}$

$\phantom{SPACER}$

$\phantom{SPACER}$

\input{figures/figure_acreage_bars2017}

\input{figures/figure_pge_ca_map}

\input{figures/figure_marg_price_5_default_rates}

\input{figures/figure_pump_hist.tex}

\input{figures/figure_hourly_consumption_vs_price}

\input{figures/figure_water_cost_cartoons}

%\input{figures/figure_davis_line_crossing.tex}

\input{figures/figure_cf_water_tax.tex}

\FloatBarrier

\input{tables/table_elec_summary_stats.tex}

\input{tables/table_water_summary_stats.tex}

\input{tables/table_elec_regs_main.tex}

\input{tables/table_water_regs_combined.tex}

\input{tables/table_water_months_from_pump_test}

\input{tables/table_water_intens_extens.tex}

\input{tables/table_probit_results.tex}

\input{tables/table_sim_acres.tex}

\end{document}
